---
title: "Data summary"
author: "Chris Beeley"
date: "25/01/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(tidylog)

if(!file.exists("all_data.Rdata")){
  
  la_data <- read_excel("LA-Type-B-November-2019-fZk64.xls", sheet = 4, skip = 13)
  
  # take out the top row
  
  la_data <- la_data %>% 
    slice(-1)
  
  trust_data <- read_excel("Trust-Type-B-November-2019-fZk64.xls", sheet = 4, skip = 13)
  
  # take out the top two rows
  
  trust_data <- trust_data %>% 
    slice(c(-1, -2))
  
  # take out the top row
  
  granular_data <- read_csv("MSitDT-NOVEMBER-2019-full-extract-for-publication-fZk64.csv")
  
  save(la_data, trust_data, granular_data, file = "all_data.Rdata")
  
} else {
  
  load("all_data.Rdata")
}

```

## Data summary

The data is published [here](https://www.england.nhs.uk/statistics/statistical-work-areas/delayed-transfers-of-care/delayed-transfers-of-care-data-2019-20/). There are three different types of data:

* Time series for England. This does not look very useful since it is very high level
* Total Delayed Days Local Authority. High level summary of delays by local authority
* Total Delayed Days Trust. High level summary of delays by health Trust
* CSV Format Monthly Delayed Transfers of Care. Granular data given by 

### Total delayed days by local authority

There are several sheets here. I've read sheet 4 for now which gives the reason for the DTOC.

### Total delayed days by Trust

There are again several sheets here. I've read sheet 4 for now which gives the reason for the DTOC.

## Shiny application

Toolkit is [here](https://improvement.nhs.uk/resources/delayed-transfer-care-dtoc-improvement-tool/). It contains:

* Pareto graph of reason for discharge
* SPC of DTOCs (all reason)
* And then SPC by reason

### Elements of Shiny application

#### Pareto graph (granular data)

```{r}

granular_data %>% 
  filter(`Provider Parent Name` == "NHS ENGLAND EAST OF ENGLAND (EAST)") %>% 
  group_by(`Reason For Delay`) %>% 
  summarise(nhs_dtoc = sum(`NHS DTOC beds`)) %>% 
  arrange(-nhs_dtoc) %>% 
  mutate(`Reason For Delay` = factor(`Reason For Delay`, levels = `Reason For Delay`)) %>% 
  mutate(csum = cumsum(nhs_dtoc)) %>% 
  ggplot(aes(x = `Reason For Delay`)) + 
  geom_bar(aes(y = nhs_dtoc), stat = "identity") + 
  geom_point(aes(y = csum)) +
  geom_path(aes(y = csum, group=1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#### SPCs of DTOCs (all reasons)
